Do expert pianists adapt their pedagogical demonstrations to novices' skills?

### 2022-11-29 paper/manuscript.tex
update results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={How do expert pianists adapt their pedagogical demonstration to novices' skill levels?},
  pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{How do expert pianists adapt their pedagogical demonstration to novices' skill levels?}
\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
\date{}

\shorttitle{How do expert pianists adapt their pedagogical demonstration to novices' skill levels?}

@@

Teaching is effective when experts and novices dynamically interact with each other (Byrne \& Rapaport, 2011). In the literature on cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they have acquired to the next generation (Mesoudi, 2016). However, information flows between experts and novices are mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022). Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).

Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019). When experts have an intention to teach, they modulate their behaviour. For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002). Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).

@@

We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference). The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA. For post-hoc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.

@@

Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) or Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002). Therefore, participants did not change the tempo depending on the types of recordings (\emph{Fig \ref{fig:ioi-1}}).

@@

There was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002). Participants produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).

@@

There was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.093, \(\eta_G^2\) = 0.001). Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).

@@

Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) or Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000). Participants did not change performances in terms of notated forte depending on the types of recordings (\emph{Fig \ref{fig:vel-1}}, left).

@@

There was a significant main effect of Articulation (\emph{F}(1, 19) = 7.18, \emph{p} = 0.01, \(\eta_G^2\) = 0.060). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000). Participants produced softer piano when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).

@@

Neither main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) or Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.336, \(\eta_G^2\) = 0.001).

@@

Neither main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.150, \(\eta_G^2\) = 0.005) or Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).

These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of recordings (\emph{Fig \ref{fig:vel-diff-1}}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
\end{figure}

@@

\hypertarget{comparison-with-baseline-performance}{%
\subsection{Comparison with baseline performance}\label{comparison-with-baseline-performance}}

In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced before listening to any of the recordings), we performed one-way repeated-measures ANOVAs. For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent. To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer. By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.

We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of novices' skill levels. We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.

There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (\emph{Fig \ref{fig:ioi-2}}).

We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not in each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.

There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.060; Greenhouse-Geisser corrected). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).

There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, right).

We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.

There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, left).

There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).

We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.

These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-diff-2}}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
\end{figure}

@@

\leavevmode\vadjust pre{\hypertarget{ref-mesoudi_2016a}{}}%
Mesoudi, A. (2016). Cultural evolution: Integrating psychology, evolution and culture. \emph{Current Opinion in Psychology}, \emph{7}, 17--22. \url{https://doi.org/10.1016/j.copsyc.2015.07.001}

@@

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Do expert pianists adapt their pedagogical demonstrations to novices' skills?},
  pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Do expert pianists adapt their pedagogical demonstrations to novices' skills?}
\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
\date{}

\shorttitle{Do expert pianists adapt their pedagogical demonstrations to novices' skills?}

@@

Effective teaching involves experts and novices dynamically interacting with each other (Byrne \& Rapaport, 2011). Typically, information flows between experts and novices is mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022). Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).

Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts. For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002). Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).

@@

First, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference). The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA. For post-hoc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.

In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced for non-teaching purposes), we performed one-way repeated-measures ANOVAs. For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent (i.e., Legato, Staccato, Forte, Piano).

@@

Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002). This indicates that participants did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).

@@

As predicted, there was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002). There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002). Participants produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).

@@

In line with our prediction, there was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009). There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.093, \(\eta_G^2\) = 0.001). Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).

@@

Neither the main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) nor the main effect of Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000). Participants did not change their performances in terms of notated forte depending on the type of recordings (\emph{Fig \ref{fig:vel-1}}, left).

@@

There was a significant main effect of Articulation (\emph{F}(1, 19) = 7.18, \emph{p} = 0.01, \(\eta_G^2\) = 0.060). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000). Participants produced softer piano when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).

@@

Neither the main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) nor Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.336, \(\eta_G^2\) = 0.001).

@@

Neither the main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.150, \(\eta_G^2\) = 0.005) nor Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).

These results indicated that participants did not change their performances in terms of dynamics contrast depending on the type of recordings (\emph{Fig \ref{fig:vel-diff-1}}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms). A dashed line represents the tempo given by a metronome. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
\end{figure}

@@

@@

@@

\hypertarget{supplementary-material}{%
\section{Supplementary Material}\label{supplementary-material}}

\hypertarget{comparison-with-baseline-performance}{%
\subsection{2. Comparison with baseline performance}\label{comparison-with-baseline-performance}}

To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and to do their best as a performer. By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.

We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented. The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking. We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.

There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (\emph{Fig \ref{fig:ioi-2}}).

We compared the baseline performance with the performances in response to recordings where dynamics was implemented to examine whether participants performed differently depending on whether articulation was implemented or not in each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.

There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.060; Greenhouse-Geisser corrected). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001, baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had the intention to teach (\emph{Fig \ref{fig:kot-2}}, left).

There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had the intention to teach (\emph{Fig \ref{fig:kot-2}}, right).

We compared the baseline performance with the performances in response to recordings where articulation was implemented to examine whether participants performed differently depending on whether dynamics was implemented or not in each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.

There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had the intention to teach (\emph{Fig \ref{fig:vel-2}}, left).

There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had the intention to teach (\emph{Fig \ref{fig:vel-2}}, right).

We compared the baseline performance with the performances in response to recordings where articulation was implemented to examine whether participants performed differently depending on whether dynamics was implemented or not in each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.

These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had the intention to teach (\emph{Fig \ref{fig:vel-diff-2}}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms). A dashed line represents the tempo given by a metronome. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
\end{figure}
